---
name: Mansi Pravin Thanki (002128043)
title: "R Notebook"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


```{r}
library(XML)
library(RSQLite)
library(DBI)
library(knitr)
library(sqldf)
```

```{r}

fpn = "Books-v3.xml"
dbfn <- "Books.db"
dbcon <- dbConnect(RSQLite::SQLite(), dbfn)
#dbfpn = paste0(path, dbfn)
```

<!--

##############################################################################################

Q1: ERD Diagram
-->

[ERD](https://imgur.com/a/fkuQOwE)
<div style="width: 640px; height: 480px; margin: 10px; position: relative;"><iframe allowfullscreen frameborder="0" style="width:640px; height:480px" src="https://lucid.app/documents/embedded/fd6a72e0-bbb0-46bd-ab9a-2b8136ad9ed7" id="55tC6xRlXmEk"></iframe></div>


<blockquote class="imgur-embed-pub" lang="en" data-id="a/fkuQOwE" data-context="false" ><a href="//imgur.com/a/fkuQOwE"></a></blockquote><script async src="//s.imgur.com/min/embed.js" charset="utf-8"></script>

```{r}
#knitr::include_graphics(<blockquote class="imgur-embed-pub" lang="en" data-id="a/fkuQOwE" data-context="false" ><a href="//imgur.com/a/fkuQOwE"></a></blockquote><script async src="//s.imgur.com/min/embed.js" charset="utf-8"></script>)

#knitr::include_graphics(<blockquote class="imgur-embed-pub" lang="en" data-id="a/fkuQOwE" data-context="false" ><a href="//imgur.com/a/fkuQOwE"></a></blockquote><script async src="//s.imgur.com/min/embed.js" charset="utf-8"></script>)

#![](https://imgur.com/a/fkuQOwE)
```


```{r}
# Reading the XML file and parse into DOM
xmlDOM <- xmlParse(file = fpn)

# get the root node of the DOM tree
r <- xmlRoot(xmlDOM)

```

```{r}
# get number of children of root 
numbooks <- xmlSize(r)
```


<!--
##############################################################################################
Q2 Create a SQLite database that implements the schema
-->

```{sql connection=dbcon}
DROP TABLE IF EXISTS Book
```

```{sql connection=dbcon}
CREATE TABLE Book(
bookid INT PRIMARY KEY,
title TEXT,
genre TEXT,
price TEXT,
edition INT,
publish_date TEXT,
description TEXT
)
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS Author
```

```{sql connection=dbcon}
CREATE TABLE Author(
authorid INT PRIMARY KEY,
authorName TEXT
)
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS BookAuthorJunction
```

```{sql connection=dbcon}
CREATE TABLE BookAuthorJunction(
bookid INT,
authorid INT,
PRIMARY KEY (bookid,authorid)
FOREIGN KEY (bookid) REFERENCES Book(bookid)
FOREIGN KEY (authorid) REFERENCES Book(authorid)
)

```

<!-- 
##############################################################################################

Q3 LOAD DATA

-->

```{r}
books <- data.frame(bookid = vector(mode = "integer", length = numbooks),
                      title = vector(mode = "character", length = numbooks),
                      edition = vector(mode = "integer", length = numbooks),
                      genre = vector(mode = "character", length = numbooks),
                      price = vector(mode = "character", length = numbooks),
                      publish_date = vector(mode = "character", length = numbooks),
                      description = vector(mode = "character", length = numbooks))



author <- data.frame(authorid = vector(mode = "integer"),
                      authorName = vector(mode = "character")
                      )

authorBook_junction <- data.frame(authorid = vector(mode = "integer"),
                      bookid = vector(mode = "integer")
                      )
commondata <- data.frame(bookid = vector(mode = "integer", length = numbooks),
                         authorFn = vector(mode = "character", length=numbooks),
                      title = vector(mode = "character", length = numbooks),
                      edition = vector(mode = "integer", length = numbooks),
                      genre = vector(mode = "character", length = numbooks),
                      price = vector(mode = "character", length = numbooks),
                      publish_date = vector(mode = "character", length = numbooks),
                      description = vector(mode = "character", length = numbooks))

```




```{r}
for(i in 1:numbooks)
{
  abook <- r[[i]]
  
  id <- xmlAttrs(abook)
  
  title <- xmlValue(abook[[2]])
  if(xmlName(abook[[3]]) == "edition")
  {
    edition <- xmlValue(abook[[3]])
    genre <- xmlValue(abook[[4]])
    price <- xmlValue(abook[[5]])
    publish_date <- xmlValue(abook[[6]])
    description <- xmlValue(abook[[7]])
  }
  else
  {
    edition <- NULL
    genre <- xmlValue(abook[[3]])
    price <- xmlValue(abook[[4]])
    publish_date <- xmlValue(abook[[5]])
    description <- xmlValue(abook[[6]])}
  
  books[i,]$bookid <- id
  books[i,]$edition <- edition
  books[i,]$title <- title
  books[i,]$genre <- genre
  books[i,]$price <- price
  books[i,]$publish_date <- publish_date
  books[i,]$description <- description
}

books

```

# author table dataframe
```{r}
for(i in 1:numbooks)
{
  auth <- r[[i]]
  id <- xmlAttrs(auth)
  author_name <- xmlValue(auth[[1]])
  
  if (!(author_name %in% author$authorName)){
  author[i,]$authorid <- i
  
  author[i,]$authorName <- author_name
  }
  
}
author_table <-na.omit(author)
author_table

```

# common data frame
```{r}
for(i in 1:numbooks)
{
  abook <- r[[i]]
  
  id <- xmlAttrs(abook)
  author <- xmlValue(abook[[1]])
  title <- xmlValue(abook[[2]])
  if(xmlName(abook[[3]]) == "edition")
  {
    edition <- xmlValue(abook[[3]])
    genre <- xmlValue(abook[[4]])
    price <- xmlValue(abook[[5]])
    publish_date <- xmlValue(abook[[6]])
    description <- xmlValue(abook[[7]])
  }
  else
  {
    edition <- NULL
    genre <- xmlValue(abook[[3]])
    price <- xmlValue(abook[[4]])
    publish_date <- xmlValue(abook[[5]])
    description <- xmlValue(abook[[6]])}
  
  commondata[i,]$bookid <- id
  commondata[i,]$authorFn <- author
  commondata[i,]$edition <- edition
  commondata[i,]$title <- title
  commondata[i,]$genre <- genre
  commondata[i,]$price <- price
  commondata[i,]$publish_date <- publish_date
  commondata[i,]$description <- description
}
#as1$Values[match(as2$ID, as1$ID)] <- as2$Values
commondata

```
# dataframe for junction table
```{r connection =dbcon}
df <- sqldf(c("select bookid, A.authorid
    FROM commondata c
    INNER JOIN author_table A ON 
        c.authorFn = A.authorName"))

df
```

```{r}
commondata
```

<!--  
##############################################################################################

Q4 Transform data types as necessary and then save the data frames to the appropriate tables in the database using a strategy of your choice
-->

# using dbWriteTable for inserting data into Book, Author and BookAuthorJunction table
```{r}
dbWriteTable(dbcon,"Book",books,overwrite=T)
```

# Book table in database
```{sql connection=dbcon}
select * from Book
```
```{r}
dbWriteTable(dbcon,"Author",author_table,overwrite=T)
```

```{sql connection=dbcon}
select * from Author
```

```{r}
dbWriteTable(dbcon,"BookAuthorJunction",df,overwrite=T)
```

```{sql connection=dbcon}
select * from BookAuthorJunction
```
<!--
##############################################################################################
Q5
-->

# Q5.A
# What is the number of genres have at least three books?
```{sql connection=dbcon}
SELECT COUNT(*)FROM (SELECT genre FROM Book 
GROUP BY genre 
HAVING count(bookid) >= 3)
```

# Q5.B
# What is the oldest year in which a publication was published?
```{sql connection=dbcon}
SELECT MAX(strftime('%Y', publish_date)) AS OLDEST_BOOK_PUBLISH_YEAR 
FROM book
```

# Q5.C
# Find the number of books and average price for each genre.
```{sql connection=dbcon}
SELECT genre, COUNT(*) AS BookCount, AVG(price) AS Avg_Price
FROM Book 
GROUP BY genre
```

# Q5.D
# List the title and author of all books that are less than 0.8*AVG or more than 1.2*AVG, where AVG is the average price of all books. As a predicate logic expression: { b : books(b) & (b.price < 0.8*AVG(books.price) | b.price > 1.2*AVG(books.price)) }

```{sql connection=dbcon}
SELECT b.title AS Title, a.authorName AS Author
FROM book B inner join
bookauthorjunction baj on b.bookid = baj.bookid
inner join author a on a.authorid = baj.authorid
WHERE (price > (SELECT AVG(price)*1.2 FROM book)) OR (price < (SELECT AVG(price)*0.8 FROM book))
```

```{r}
dbDisconnect(dbcon)
```

