---
title: "Indexing Assignment"
output: html_notebook
Author1: Mansi Pravin Thanki (thanki.m@northeastern.edu)
Author2: Hardik Nahata (nahata.h@northeastern.edu)

---

<!-- ################################ Experiments with SQLite ############################################# -->
# Experiments for MYSQL starts at Line 302 of this notebook 

```{r}
library(RSQLite)
file_path = (getwd())
dbfile = "/sakila.db"
dbcon <- dbConnect(RSQLite::SQLite(), paste0(file_path,dbfile))
```


<!-- 
Q1: query on a single table with a WHERE clause that is not the primary key but has an index; 
same query but no index on that column 
-->

<!--
          1. Table used: film table
          2. key used in where clause: rating -> not a primary key
          On running the EXPLAIN QUERY Plan:
          we can observe that -> detail: SCAN film 
          since a non-primary key is used in where clause for search, SCAN film is displayed.
          
          This denotes that the table was scanned and the index is not used
          
          This observation matches the information in:                                                                                         http://artificium.us/lessons/60.dbdesign/l-60-611-indexing-sqlite/l-60-611.html
-->

```{sql connection=dbcon}

EXPLAIN QUERY PLAN
SELECT * 
FROM film 
WHERE rating = 'R'

```

<!-- Experiment with query execution time without indexing -->

<!--
      Using sys.Time() for noting the start and end time stamp
      Index has not been created on rating column
      Time for execution can be seen on the R console
-->

```{r connection = dbcon}
start_time <- Sys.time()

dbGetQuery(dbcon,
  "select * from film where rating = 'R';")


end_time <- Sys.time()

print("For Q1, execution speed for non-indexed version:")
end_time - start_time

```



# creating an index on film table on the column of rating
```{sql connection=dbcon}
CREATE INDEX filmRating 
    ON film (rating);
```


<!--
          
          On running the EXPLAIN QUERY Plan:
          we can observe that -> detail: SEARCH film USING INDEX filmRating (rating=?)
          since a non-primary key is used in where clause for search, SCAN film is displayed.
          
          This denotes that the table was searched and the index is used
          
          This observation matches the information in:                                                                                         http://artificium.us/lessons/60.dbdesign/l-60-611-indexing-sqlite/l-60-611.html
-->

```{sql connection=dbcon}
EXPLAIN QUERY PLAN
select * from film where rating = 'R'

```

<!--
      Using sys.Time() for noting the start and end time stamp
      Index has now been created on rating column
      Time for execution can be seen on the R console
      
      It can be observed that the EXECUTION SPEED OF INDEXED VERSION IS MUCH MORE FASTER THAN THE NON-INDEXED VERSION
-->

```{r connection=dbcon}
start_time <- Sys.time()

dbExecute(dbcon,
  "select * from film where rating = 'R';")

end_time <- Sys.time()
print("For Q1, execution speed for indexed version:")
end_time - start_time
```

# dropping the filmRating index
```{sql connection=dbcon}
drop index filmRating

```


<!---------- Q1 ENDS --------------->


<!-- 
#Q2: query with a join on at least three tables (no secondary indexes; no where clause other than join clause) 


First, we are checking the already available indexed columns in database using the below query -->

```{sql connection=dbcon}
SELECT
   `type`, 
   `name`, 
   `tbl_name`, 
   `sql`
  FROM sqlite_master
WHERE `type` = 'index';

```

<!--
          1. Table used: film_actor, film, film_category  table
         
-->


```{sql connection=dbcon}
select * from FILM_ACTOR FA inner join FILM F on FA.film_id= F.film_id inner join Film_Category FC on F.film_id=FC.film_id
```
<!--
          On running the EXPLAIN QUERY Plan:
          we can observe that -> 
          detail: 
          SCAN FA 
          SEARCH F USING INTEGER PRIMARY KEY (rowid=?)
          SEARCH FC USING INDEX sqlite_autoindex_film_category_1 (film_id=?)	

          This denotes that:
          1. Film_actor table was scanned without considering any index, and 
          2. Film,Film_Category table were searched using index
          
          This observation matches the information in:                                                                                         http://artificium.us/lessons/60.dbdesign/l-60-611-indexing-sqlite/l-60-611.html
-->


```{sql connection=dbcon}
EXPLAIN QUERY PLAN
select * from FILM_ACTOR FA inner join FILM F on FA.film_id= F.film_id inner join Film_Category FC on F.film_id=FC.film_id

```

<!---------------- Q2 Ends ---------------->


<!-- Q3: query on a single table with group by having a where clause that involves only the primary key column-->

<!--      1. Table used: customer table
          2. key used in where clause: customer_id -> is a primary key
          3. group by: active
          
-->

```{sql connection=dbcon}
SELECT *
FROM customer
WHERE customer_id >=1 AND customer_id<=100
GROUP BY active

```

<!-- 
          On running the EXPLAIN QUERY Plan:
          we can observe that -> detail: 
          1. SEARCH customer USING INTEGER PRIMARY KEY (rowid>? AND rowid<?)
          -  This denotes that the customer table was searched using the index on primary key
          
          2. USE TEMP B-TREE FOR GROUP BY:
          -  SQLite uses temporary B+ trees to sort the output rows when any select query contains order by, group by or distinct                 clause. Since, we have group by in the query, SQLite used the temporary B+ Tree
-->

```{sql connection=dbcon}
EXPLAIN QUERY PLAN
SELECT *
FROM customer
WHERE customer_id >=1 AND customer_id<=100
GROUP BY active
```


<!-- Q3 Ends --->

<!-- Q4: one additional experiment of your choice: any query, any index use -->

<!-- 
SORTING: 
We are conducting experiment on the execution speed when index is involved and when index is not involved during sorting.

          1. Table used: actor table
          2. key used in where clause: actor_id 
          3. ordering in descending order of first_name
          
-->

```{sql connection=dbcon}
select * from actor
   where actor_id between 1 and 100
   order by first_name desc;
   
```

<!-- 
          On running the EXPLAIN QUERY Plan:
          we can observe that -> detail: 
          1. SCAN actor
          -  This denotes that the actor table was scanned and index was not considered 
          
          2. USE TEMP B-TREE FOR ORDER BY:
          -  SQLite uses temporary B+ trees to sort the output rows when any select query contains order by, group by or distinct                 clause. Since, we have group by in the query, SQLite used the temporary B+ Tree
-->

```{sql connection=dbcon}
explain query plan
select * from actor
order by first_name desc
   
```

<!--
      Using sys.Time() for noting the start and end time stamp
      Index has not been created on first_name column
      Time for execution can be seen on the R console
-->

```{r connection=dbcon}
start_time <- Sys.time()

dbExecute(dbcon,
  "select * from actor
   where actor_id between 1 and 100
   order by first_name desc;")

end_time <- Sys.time()
print("For Q4, execution speed for non-indexed version:")
end_time - start_time
```

# creating index on the first_name as fName

```{sql connection=dbcon}
CREATE INDEX fName 
    ON actor (first_name);
```

<!--
      Using sys.Time() for noting the start and end time stamp
      Index has now been created on first_name column
      Time for execution can be seen on the R console
      
      It can be observed that the EXECUTION SPEED OF INDEXED VERSION IS SLIGHTLY FASTER THAN THE NON-INDEXED VERSION
-->

```{r connection=dbcon}
start_time <- Sys.time()

dbExecute(dbcon,
  "select * from actor
   where actor_id between 1 and 100
   order by first_name desc;")

end_time <- Sys.time()
print("For Q4, execution speed for indexed version:")
end_time - start_time
```
# droping the created index

```{sql connection=dbcon}
drop index if exists fName
```

# disconnecting dbcon
```{r}
dbDisconnect(dbcon)

```

<!-- ################################ Experiments with MYSQL ############################################# -->

```{r}
rm(list = ls())
library(DBI)
library(RMySQL)
library(sqldf)
mysqlcon<-dbConnect(MySQL(),host="localhost",port=3306,dbname='sakila',user="root",password="root12345")

```

<!-- 
Q1: query on a single table with a WHERE clause that is not the primary key but has an index; 
same query but no index on that column 
-->

<!--
          1. Table used: film table
          2. key used in where clause: rating -> not a primary key
          3. Execution speed for non-indexed rating version is printed in the R Console
-->

```{r connection = mysqlcon}

start_time <- Sys.time()

suppressWarnings(dbGetQuery(mysqlcon,
  "select * from film where rating = 'R';"))
end_time <- Sys.time()
print("For Q1, (MYSQL) execution speed for non-indexed version:")
print(end_time - start_time)
```

<!--
          
          On running the query EXPLAIN:
          we can observe that -> 
          Since there are no keys, possible_keys and key and key_len are empty.
          It denotes that no relevant index exists were found on the table
          
-->


```{sql connection=mysqlcon}
explain
select * from film where rating = 'R';
```


# creating index on rating as filmRating

```{sql connection=mysqlcon}
CREATE INDEX filmRating 
    ON film (rating);
```

<!--
          
          On running the query EXPLAIN after creating index:
          we can observe that -> 
          Since we created index on rating, the  keys, possible_keys and key and key_len are filled with 
          filmRating values and correct key length.
          
          Extra column also now denotes that index condition is being used
          
          
-->

```{sql connection=mysqlcon}
EXPLAIN 
select * from film where rating = 'R'

```

<!--
          
          Execution speed for indexed rating version is printed in the R Console
-->

```{r connection = mysqlcon}

start_time <- Sys.time()

suppressWarnings(dbGetQuery(mysqlcon,
  "select * from film where rating = 'R';"))
end_time <- Sys.time()
print("For Q1, (MYSQL) execution speed for indexed version:")
print(end_time - start_time)
```

# dropping filmRating index
```{sql connection=mysqlcon}
drop index filmRating on film
```

<!---- Q1 ENDS----->

<!--
Q2: query with a join on at least three tables (no secondary indexes; no where clause other than join clause) 

1. Tables used: film_actor, film, film_category table

-->


```{sql connection=mysqlcon}
select * from FILM_ACTOR FA inner join FILM F on FA.film_id= F.film_id inner join Film_Category FC on F.film_id=FC.film_id
```

<!--
          On running the query EXPLAIN :
          we can observe that -> 
          
          possible_keys: show the possible keys used for indexing in all three tables
          
          key: shows the actual key that was used by MySql for indexing
          key values can be seen for Film and Film_Actor table

          This denotes that:
          1. Film_Category table was scanned without considering any index, and 
          2. Film,Film_Actor table were searched using index
          
          Unlike in experiment for SQLite, where FA was scanned without using index, here index is used.
-->

```{sql connection=mysqlcon}
EXPLAIN 
select * from FILM_ACTOR FA inner join FILM F on FA.film_id= F.film_id inner join Film_Category FC on F.film_id=FC.film_id

```

<!--
          Using sys.Time() for noting the start and end time stamp
          Time for execution can be seen on the R console
-->
```{r connection = mysqlcon}

start_time <- Sys.time()

suppressWarnings(dbGetQuery(mysqlcon,
  "select * from FILM_ACTOR FA inner join FILM F on FA.film_id= F.film_id inner join Film_Category FC on F.film_id=FC.film_id"))
end_time <- Sys.time()
print("For Q2, MySQl version execution speed:")
print(end_time - start_time)
```


<!------ Q2 ENDS ----->


<!-- Q3: query on a single table with group by having a where clause that involves only the primary key column-->

<!--      1. Table used: customer table
          2. key used in where clause: customer_id -> is a primary key
          3. group by: active
          
-->

```{sql connection=mysqlcon}
SELECT *
FROM customer
WHERE customer_id >=1 AND customer_id<=100
GROUP BY active
```

<!-- 
          On running the query EXPLAIN:
          we can observe that:
          key: contains PRIMARY
          -  This denotes that the customer table was searched using the index on primary key
          
          In Extra column: Using where; Using temporary
          - This denotes that MySQL used temporary table in memory for the given query
-->


```{sql connection=mysqlcon}
EXPLAIN 
SELECT *
FROM customer
WHERE customer_id BETWEEN 1 AND 100
GROUP BY active
```

<!--
          Using sys.Time() for noting the start and end time stamp
          Time for execution can be seen on the R console
-->

```{r connection=mysqlcon}
start_time <- Sys.time()

suppressWarnings(dbGetQuery(mysqlcon,
  "SELECT *
FROM customer
WHERE customer_id BETWEEN 1 AND 100
GROUP BY active"))
end_time <- Sys.time()
print("For Q2, MYSQL version execution speed:")
print(end_time - start_time)
```

<!-- Q3 Ends --->

<!-- Q4: one additional experiment of your choice: any query, any index use -->

<!-- 
SORTING: 
We are conducting experiment on the execution speed when index is involved and when index is not involved during sorting.

          1. Table used: actor table
          2. key used in where clause: actor_id
          3. ordering in descending order of first_name

          Using sys.Time() for noting the start and end time stamp
          Index has now been created on first_name column
          Time for execution can be seen on the R console
-->
```{r connection=mysqlcon}
start_time <- Sys.time()

suppressWarnings(dbGetQuery(mysqlcon,
  "select * from actor
   where actor_id between 1 and 100
   order by first_name desc;"))
end_time <- Sys.time()
print("For Q4 (Mysql), execution speed for non-indexed version")
print(end_time - start_time)
```

<!-- 
          On running the EXPLAIN QUERY Plan:
          we can observe that -> detail: 
          key: contains primary
          -  This denotes that the actor table was searched using the index 
          
          In Extra column:
          - This denotes that MySQL used filesort for the ordering of the files
-->

```{sql connection=mysqlcon}
EXPLAIN
select * from actor
   where actor_id between 1 and 100
   order by first_name desc;
```

# creating index on first_name column of actor as fName
```{sql connection=mysqlcon}
CREATE  INDEX fName 
    ON actor (first_name);
```

<!--
      Using sys.Time() for noting the start and end time stamp
      Index has now been created on first_name column
      Time for execution can be seen on the R console
      
      It can be observed that the EXECUTION SPEED OF INDEXED VERSION IS MUCH MORE FASTER THAN THE NON-INDEXED VERSION
-->

```{r connection=mysqlcon}
start_time <- Sys.time()

suppressWarnings(dbGetQuery(mysqlcon,
  "select * from actor
   where actor_id between 1 and 100
   order by first_name desc;"))
end_time <- Sys.time()
print("For Q4 (Mysql), execution speed for indexed version")
print(end_time - start_time)
```
# dropping the created index
```{sql connection=mysqlcon}
drop index fName on actor
```


<!---

Observations from all experiments:

1. Indexing improved the speed of query execution in both SQLite and MYSQL

2. MYSQL had a better performance in terms of speed as compared to SQLite while considering joins on 3 tables

3. Indexing on primary key displayed faster execution

4. For Joins, sqlite and mysql considered different table for indexing 

-->

# disconnecting from mysqlcon
```{r connection=mysqlcon}
dbDisconnect(mysqlcon)
```